---
title: "Loading-Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Loading-Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GDSCtools)
library(purrr)
```

# Introduction

This vignette demonstrates how to load in multiple samples from `CellRanger` output folders as `Seurat` objects using `GDSCtools`

The workflow includes:

1. Loading in `CellRanger` out files as `Seurat` objects. 

First, designate the path to the `CellRanger` files, the sample names stored within that folder, and the corresponding IDs you wish to name the samples.

`purrr::map` is used to apply the `GDSCtools::read_10x_GEX` function to all the samples in the input directory, returning a list of Seurat objects. The resulting objects are pre-filtered based on `min.cells` and `min.features`.

```{r eval=FALSE}

sample_dir <- "/path/to/data/"
samples    <- c("S1", "S2", "S3")
idents     <- c("CondA", "CondB", "CondC")

seurat_list <- map(samples, idents, ~ read_10x_GEX(
                   sampleDir    = sample_dir,
                   sample       = .x,
                   ident        = .y,
                   min.cells    = 10,
                   min.features = 100))

```

