---
title: "GO-ORA Network Analysis Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GO-ORA Network Analysis Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

# Introduction

This vignette demonstrates a typical workflow for differential expression,
GO over-representation analysis (GO-ORA), pathway overlap network generation,
and pathway module interpretation using `GDSCtools`.

The workflow includes:

1. Differential expression analysis
2. GO enrichment analysis
3. Pathway overlap network construction
4. Leiden clustering of GO pathways
5. Word cloud generation
6. Extraction of biologically related GO modules

# Load Packages

```{r eval = FALSE}
library(GDSCtools)
library(org.Mm.eg.db)
```

# Input Data

This workflow assumes a Seurat object named `seurat_obj`
containing single-cell RNA-seq data.

The identities being compared in this example are:

- `Car`
- `Bleo`

across selected fibroblast clusters.

# Differential Expression Analysis

Run differential expression analysis across selected clusters.

```{r eval = FALSE}
de_results <- run_cluster_de(
  obj        = seurat_obj,
  clusters   = c("Fibroblast 1", "Fibroblast 2"),
  ident1     = "Car",
  ident2     = "Bleo"
)
```

Write differential expression results to disk.

```{r eval = FALSE}
write_de_results(
  de_results,
  outSubDir = "results/DE/",
  type = "DE"
)
```

# GO Over-Representation Analysis

Define the gene universe.

```{r eval = FALSE}
universe <- rownames(seurat_obj)
```

Run GO enrichment analysis using Biological Process ontology.

```{r eval = FALSE}
go_results <- run_go_ora(
  de_list   = de_results,
  org_db    = org.Mm.eg.db,
  universe  = universe,
  ont       = "BP",
  outputDir = "results/GO/"
)
```

Export GO enrichment results.

```{r eval = FALSE}
write_de_results(
  go_results,
  outSubDir = "results/GO/",
  type = "GO",
  split_direction = FALSE
)
```

# Pathway Overlap Network Analysis

Generate pathway overlap networks from GO enrichment results.

```{r eval = FALSE}
net_results <- run_comparison_networks(
  go_list     = go_results,
  de_list     = de_results,
  comp_label  = "Car_vs_Bleo",
  base_outdir = "results/"
)
```

The resulting networks:

- represent GO terms as nodes,
- connect pathways sharing DE genes,
- cluster pathways using Leiden community detection,
- and identify higher-order biological programs.

# Export GO Cluster Membership

Export GO terms and their associated Leiden clusters.

```{r eval = FALSE}
for (i in names(net_results)) {

  x <- net_results[[i]]$go_term_df

  cluster <- gsub("/", "_", i)

  name <- paste0(cluster, "_cluster_membership.csv")

  write.csv(
    x,
    file = paste0(
      outputDir,
      "CD206_Car_vs_Bleo/networks/",
      name
    )
  )
}
```

# Generate Word Clouds

Generate cluster-specific pathway word clouds.

```{r eval = FALSE}
write_cluster_wordclouds_patchwork(
  net_list   = net_results,
  outputDir  = "results/",
  comparison = "Car_vs_Bleo"
)
```

These word clouds summarize dominant biological themes
within each Leiden pathway module.

# Inspect a Specific Biological Module

Extract all GO terms belonging to the same Leiden cluster
as a target GO term.

For example, the extracellular matrix organization module:

```{r eval = FALSE}
ecm_terms <- write_cluster_results(
  net_result = net_results[["Fibroblast_1"]],
  go_term    = "extracellular matrix organization",
  outdir     = "results/Car_vs_Bleo/networks/"
)
```

This can help identify coordinated biological programs such as:

- fibrosis,
- inflammatory signaling,
- chemotaxis,
- interferon response,
- or metabolic remodeling.

# Session Information

```{r}
sessionInfo()
```
